Temporal Continuity and the Judgment of Actual Causation

### 2021-01-27 docs/resources/CogSci_paper/paper.tex
paper modified
added stats

abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract 
abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract abstract

\textbf{Keywords: causal inference; causal selection; causal explanation} 
\end{abstract}

\section{Introduction}

Around 5:27 pm on November 9, 1965, Martin Saltzman found himself trapped in a dark elevator roughly a quarter of a way up the Empire State building~\cite{gelbr1965night}. The stopping of Saltzman's elevator was one of thousands of unexpected events that occurred that evening, and virtually all of the people involved must have wondered about the causes of these events. Subsequent accounts of the Great Northeastern Blackout often trace these events back to the tripping of a relay on line Q29BD in Ontario that triggered a cascade of failures.

Identifying the cause of an event (e.g.\ the stopping of an elevator) requires a judgment of actual causation, also known as singular or token causation~\cite{danks17}. Psychologists and philosophers have explored these judgments in detail~and developed formal models of actual causation, including models based on Bayesian networks \cite{halpern16}, force dynamics \cite{wolfft17} and mental simulation~\cite{gerstenberg20}.  Here we present and evaluate an account of actual causation that highlights the role of temporal continuity.  On this account, the tripping of relay Q29BD is the main cause of the elevator stopping because the tripping initiated a continuous sequence of state changes that culminated in the stopping of the elevator.

@@

The continuity account relies on two core principles. First, it assumes that the effect to be explained and its cause are both changes of state. The tripping of a relay qualifies as a candidate cause, but the steady-state setting of a relay does not. Our emphasis on state changes is consistent with the common view that causal relationships are relationships between events rather than facts or states of affairs, but  \citeA{glymourstone} point out that most Bayes net accounts of actual causation ignore changes of state. State changes are embedded in continuous time, and a change that begins at a specific moment is typically the direct result of an event that occurred an instant before. Focusing on state changes therefore motivates a principle of temporal continuity that allows continuous causal sequences to be traced back in time from an effect to its main cause. Some previous work on actual causation highlights the importance of time~\cite{youngs09,stephanmw20e}, but to our knowledge previous work has not explored the implications of temporal continuity in the way we do here.

@@

\begin{figure*}
\begin{center}
%\includegraphics[width=7cm]{figures/demo.eps}
\includegraphics{figures/introfig.eps}
\end{center}
\caption{(a)(i) A smoke alarm connected to a chain of ionization detectors (top branch) and a chain of photoelectric detectors (bottom branch). All components are initially inactive (ii) Detector \ev{R}{D} activates and activity propagates along the top branch. (iii) Detector \ev{R}{C} activates and ultimately triggers the activation of the alarm, which activates only when both  \ev{I}{D} and \ev{I}{C} are active. (b) An activation sequence in which (i) the top branch activates first, (ii) activation starts and finishes along the bottom branch while activation continues along the top branch, and (iii) the top branch completes its activation and ultimately triggers the smoke alarm.}
\label{fig:intro}
\end{figure*}

@@

The boundaries of the causal system under consideration will depend on context. For simplicity, our discussion of Figure~\ref{fig:intro}a has focused only on events internal to the particle detection network, and has attemped to characterize which of these events is best viewed as the main cause of the GCM's activation. The particle detection network, however, could also be considered part of a broader causal system that includes both the network and the network's surroundings. For example, if the network is embedded in an apparatus for carbon dating, then the spontaneous decay of a carbon atom will probably be identified as the immediate cause of \ev{R}{C} and the main cause of the GCM's activation.

@@

\textbf{Participants}. 30 participants were recruited via Amazon Mechanical Turk and paid XXX for a YYY minute experiment.

@@

\textbf{Materials}. The experiment used a customized interface built using the jsPsych library~\cite{XXX}. For all networks presented, participants clicked a ``Run'' button to observe an activation sequence. The first event in the sequence (i.e.\ the first change of state) always took place 5 seconds after the Run button was clicked, and the delay between successive activations along a chain of detectors was set to 100 ms. After the final event in a sequence (i.e.\ the activation of the GCM), all detectors became clickable after a delay of 1 second. Clicking on a detector turned its border red, and at most one detector could be selected at any time. After a sequence completed, participants could view it again if they wished by clicking a ``Run again'' button.

@@

Each dual branch sequence (short, medium and long) came in four versions. The \emph{state} version showed the delayed branch (including detectors \ev{R}{D} and \ev{I}{D}) as active from the very beginning of the sequence. The activation of this branch was therefore presented as having occurred at some indefinite time in the past, resulting in a steady state of activation. The three \emph{event} versions all showed \ev{R}{D} activating 5 seconds into the sequence, and had delays of 2, 4 and 6 seconds between the activation of \ev{I}{D} and \ev{R}{C}.

@@

Although \ev{R}{D}, \ev{I}{D}, \ev{R}{C} and \ev{I}{C} all qualify as causes of the effect, Figure~\ref{fig:expt1} suggests that \ev{R}{C} tends to be singled out as the main cause. This result can be separated into two general conclusions.  First,  participants were more likely to choose root causes than immediate causes. Root and immediate causes were identical for the short sequences and therefore cannot bet distinguished, but the results for medium and long sequences reveal a preference for root causes. STATISTICAL ANALYSIS.

The second conclusion is that for dual branch sequences, the root cause on the continuous branch (\ev{R}{C}) is preferred to the root cause on the delayed branch (\ev{R}{D}). This result is consistent with the continuity account but also consistent with the probability raising account, because \ev{R}{C} guarantees the occurrence of the effect but \ev{R}{D} does not.

Figure \ref{fig:intro} suggests that event and state sequences lead to similar patterns of responses. The continuity account makes the same prediction for both kinds of sequences, but we anticipated that the event sequences might make participants especially likely to choose \ev{R}{C} over \ev{R}{D}. If this effect exists, our data suggest that it is relatively small.

@@

Thus far we have focused on dual branch sequences, and the chain sequences are of interest primarily because they provide additional evidence of a preference for root causes.  NUMBERS HERE

@@

\begin{figure}
\begin{center}
%\includegraphics[width=8cm]{results_E1}
\includegraphics{figures/expt1.eps}
\end{center}
\caption{Experiment 1. (a) Order of the four key events for dual branch sequences. The probability raising and continuity accounts both single out \ev{R}{C} as the main cause. (b)  Responses for dual branch sequences. The delayed branch was either inactive (event sequences) or active (state sequences) at the start of the sequence.} 
\label{fig:expt1}
\end{figure}

@@

\textbf{Participants}. 100 participants were recruited via Amazon Mechanical Turk and paid XXX for a YYY minute experiment. Two did not submit a valid completion code, leaving 98 sets of responses for analysis.

@@

\textbf{Results}.  Network orientation (left or right) and position of the curved branch (top or bottom) had no effect and we collapsed across these variables.  Figure~\ref{fig:expt2}c summarizes the results for the three orders listed in Figures \ref{fig:expt2}a and \ref{fig:expt2}b. Consistent with Experiment 1,  \ev{R}{C} is preferred over \ev{R}{D} given the order \{\ev{R}{D},\ev{I}{D},\ev{R}{C},\ev{I}{C}\}, and a similar (but weaker) effect was found for the \{\ev{R}{D},\ev{R}{C},\ev{I}{D},\ev{I}{C}\} sequences. The critical finding is that \ev{R}{C} is also preferred for the \{\ev{R}{C},\ev{R}{D},\ev{I}{D},\ev{I}{C}\} sequences, even though the probability raising account makes the opposite prediction. STATISTICS HERE.

Relative to Experiment 1, the frequency with which \ev{I}{C} is chosen has increased in Experiment 2. This difference may reflect the increased difficulty of Experiment 2. When both branches of a dual branch structure are simultaneously active, keeping track of both root causes and the order in which they occurred is relatively challenging, which may lead some participants to fall back on the simple strategy fo choosing the immediate cause that directly precedes the effect.

@@

\begin{figure}
\begin{center}
%\includegraphics[width=8cm]{results_E2}
\includegraphics{figures/expt2.eps}
\end{center}
\caption{Experiment 2. (a) The probability raising and continuity accounts make different predictions for dual branch sequences in which \ev{R}{C} occurs before \ev{R}{D}. (b) The two accounts agree when \ev{R}{C} occurs after \ev{R}{D}.
(c) Responses for the three possible orders of the key events.} 
\label{fig:expt2}
\end{figure}

@@

Because the continuity account considers changes of state in continuous time, it can exploit a ``no coincidences'' principle to identify a single main cause of an effect. This principle applies broadly to causation in physical systems, but is less applicable to social scenarios such as voting scenarios~\cite{livengood13} Exceptions to the ``no coincidences'' principle are possible even for physical systems, and the continuity account can explain the illusory perception of actual causation that occurs in some such cases~\cite{thorstadw16}. For example, an 11 year old boy whacked a stick against a telephone pole at the instant at which the Great Northeastern Blackout hit his town, and ran home terrified that he had caused the power outage~\cite{gelbr1965night}.

@@

As suggested earlier, standard formulations of the physical process view treat both \ev{R}{C} and \ev{R}{D} as causes of the effect in Figure~\ref{fig:intro}. Our data are consistent with the possibility that people view \ev{R}{C} and \ev{R}{D} as equally valid causes, but pick \ev{R}{D} when forced to break the tie for reasons that may be relatively superficial. For example, perhaps \ev{R}{D} is preferred because the activation on the continuous branch is continuous both in space and time, and therefore more prototypical of a causal process than the activation along the delay branch. We have planned a future study that asks whether the \ev{R}{C} over \ev{R}{D} is fundamental or superficial using a paradigm in which participants rank multiple candidate causes in order of importance, and in which ties are permitted.

@@

>>>>

Psychological theories of actual causation aim to characterize which of multiple causes of an event is singled out as the primary cause. We present one such theory called the continuity account of actual causation. The continuity account treats events as changes of state in continuous time and traces a sequence of stage changes backwards through time from an event to its primary cause. The account is broadly compatible with the physical process view of causation, and we test it by asking people to identify the primary cause of events occurring in simple physical systems. An initial experiment confirms that root causes are more likely to be chosen as primary causes than are immediate causes. A second experiment demonstrates that root causes that have temporal continuity with the effect are preferred even when probability raising accounts would predict otherwise. The results of both experiments are consistent with the continuity account, and suggest that inferences about changes of state in continuous time may underpin an important class of actual causation judgments.

\textbf{Keywords: causal inference; causal selection; singular causation; token causation; causal explanation} 
\end{abstract}

%\section{Introduction}

Around 5:27 pm on November 9, 1965, Martin Saltzman found himself trapped in a dark elevator roughly a quarter of a way up the Empire State building~\cite{gelbr1965night}. The stopping of Saltzman's elevator was one of thousands of unexpected events that occurred that evening, and all of the people involved must have wondered about the causes of these events. Subsequent accounts of the Great Northeastern Blackout often trace these events back to the tripping of a relay on line Q29BD in Ontario that triggered a cascade of failures.

Identifying the cause of an event (e.g.\ the stopping of an elevator) requires a judgment of actual causation, also known as singular or token causation~\cite{danks17}. Psychologists and philosophers have explored these judgments in detail~and have developed formal models of actual causation, including models based on Bayesian networks \cite{halpern16}, force dynamics \cite{wolfft17} and mental simulation~\cite{gerstenberg20}.  Here we present and evaluate an account of actual causation that highlights the role of temporal continuity.  On this account, the tripping of relay Q29BD is the main cause of the elevator stopping because the tripping initiated a continuous sequence of state changes that culminated in the stopping of the elevator.

@@

The continuity account relies on two core principles. First, the effect to be explained and its cause are both changes of state. The tripping of a relay qualifies as a candidate cause, but the steady-state setting of a relay does not. Our emphasis on state changes is consistent with the common view that causal relationships are relationships between events rather than facts or states of affairs, but  \citeA{glymourstone} point out that most Bayes net accounts of actual causation ignore changes of state. State changes are embedded in continuous time, and a change that begins at a specific moment is typically the direct result of an event that occurred an instant before. Focusing on state changes therefore motivates a principle of temporal continuity that allows continuous causal sequences to be traced back in time from an effect to its main cause. Some previous work on actual causation highlights the importance of time~\cite{stephanmw20e}, and our approach is highly compatible with work by Michotte and others suggesting that temporal information is often critical for identifying causes~\cite{youngs09}.  To our knowledge, however,  previous work has not explored the implications of temporal continuity in the way that we do here.

@@

%\begin{figure*}
%\begin{center}
%\includegraphics[width=7cm]{figures/demo.eps}
%\includegraphics{figures/introfig.eps}
%\end{center}
%\caption{(a)(i) A smoke alarm connected to a chain of ionization detectors (top branch) and a chain of photoelectric detectors (bottom branch). All components are initially inactive (ii) Detector \ev{R}{D} activates and activity propagates along the top branch. (iii) Detector \ev{R}{C} activates and ultimately triggers the activation of the alarm, which activates only when both  \ev{I}{D} and \ev{I}{C} are active. (b) An activation sequence in which (i) the top branch activates first, (ii) activation starts and finishes along the bottom branch while activation continues along the top branch, and (iii) the top branch completes its activation and ultimately triggers the smoke alarm.}
%\label{fig:intro}
%\end{figure*}

\begin{figure*}
\begin{center}
\null\hfill
\subfloat[\label{subfig:1a}]{%
\includegraphics{figures/fig1a.eps}
}
\hfill
\subfloat[\label{subfig:1b}]{%
\includegraphics{figures/fig1b.eps}
}
\hfill\null
\end{center}
\caption{(a)(i) A network of particle detectors. All components are initially inactive. (ii) Detector \ev{R}{D} activates and activity propagates along the top branch. (iii) Detector \ev{R}{C} activates and ultimately triggers the activation of the GCM, which activates only when both  \ev{I}{D} and \ev{I}{C} are active. (b) An activation sequence in which (i) the top branch activates first, (ii) activation starts and finishes along the bottom branch while activation continues along the top branch, and (iii) the top branch completes its activation and ultimately triggers the GCM.}
%\caption{(a)(i) A smoke alarm connected to a chain of ionization detectors (top branch) and a chain of photoelectric detectors (bottom branch). All components are initially inactive (ii) Detector \ev{R}{D} activates and activity propagates along the top branch. (iii) Detector \ev{R}{C} activates and ultimately triggers the activation of the alarm, which activates only when both  \ev{I}{D} and \ev{I}{C} are active. (b) An activation sequence in which (i) the top branch activates first, (ii) activation starts and finishes along the bottom branch while activation continues along the top branch, and (iii) the top branch completes its activation and ultimately triggers the smoke alarm.}
\label{fig:intro}
\end{figure*}

@@

The boundaries of the causal system under consideration will depend on context. For simplicity, our discussion of Figure~\ref{fig:intro}a has focused only on events internal to the particle detection network, and has attemped to characterize which of these events is best viewed as the main cause of the GCM's activation. The particle detection network, however, could also be considered part of a broader causal system that includes both the network and the network's surroundings. For example, if the network is embedded in an apparatus for carbon dating, then the spontaneous decay of a carbon atom could be identified as the immediate cause of \ev{R}{C} and the main cause of the GCM's activation.

@@

\textbf{Participants}. 30 participants were recruited via Amazon Mechanical Turk and paid \$3 for an 18 minute experiment.

@@

\textbf{Materials}. The experiment used a customized interface built using the jsPsych library~\cite{jspsych}. For all networks presented, participants clicked a ``Run'' button to observe an activation sequence. The first event in the sequence (i.e.\ the first change of state) always took place 5 seconds after the Run button was clicked, and the delay between successive activations along a chain of detectors was set to 100 ms. After the final event in a sequence (i.e.\ the activation of the GCM), all detectors became clickable after a delay of 1 second. Clicking on a detector turned its border red, and at most one detector could be selected at any time. After a sequence completed, participants could view it again if they wished by clicking a ``Run again'' button.

@@

Each dual branch sequence (short, medium and long) came in four versions. The \emph{state} version showed the delayed branch (including detectors \ev{R}{D} and \ev{I}{D}) as active from the very beginning of the sequence. The activation of this branch was therefore presented as having occurred at some indefinite time in the past, resulting in a steady state of activation. The three \emph{event} versions all showed \ev{R}{D} activating 5 seconds into the sequence, and had delays of 2, 4 and 6 seconds between the activation of \ev{I}{D} and \ev{R}{C}. The predictions of the continuity account are unaffected by the delay, but we tested different delays just in case this variable affected people's responses. The continuity account also makes the same prediction about both state and event sequences, but we anticipated that the state sequences might make participants especially likely to choose \ev{R}{C} over \ev{R}{D}.

@@

Although \ev{R}{D}, \ev{I}{D}, \ev{R}{C} and \ev{I}{C} all qualify as causes of the effect, Figure~\ref{fig:expt1} suggests that \ev{R}{C} tends to be singled out as the main cause. This result can be separated into two general conclusions.  First,  participants were more likely to choose root causes than immediate causes. Root and immediate causes were identical for the short sequences and therefore cannot be distinguished, but the results for medium and long sequences reveal a preference for root causes.  The second conclusion is that for dual branch sequences, the root cause on the continuous branch (\ev{R}{C}) is preferred to the root cause on the delayed branch (\ev{R}{D}). This result is consistent with the continuity account but also consistent with the probability raising account, because \ev{R}{C} guarantees the occurrence of the effect but \ev{R}{D} does not.

To support these two conclusions we ran two Bayesian mixed effects regression models using the \emph{brms} package in R~\cite{brms}. For the first model, each response was coded as root or immediate, and all responses that could not be classified in this way (including all responses for short sequences) were discarded. We then ran a logistic regression in which stativity (i.e.\ state vs event) and network length (long vs medium) were included as predictors of a binary dependent variable (root vs immediate). We included a random intercept for participant to allow for statistical dependencies between multiple responses from the same participant. The posterior mean of the intercept variable was 10.54 and the 95\% credible interval was $[2.30, 26.54]$, which supports the conclusion that root causes were chosen more often than immediate causes. The credible intervals for the other coefficients both included 0 (state: -2.75, $[-10.03, 1.14]$ ; long: 1.85, $[-1.78, 7.33]$), suggesting that the preference for root causes was not strongly affected by either stativity or network length.

The second analysis was very similar except that the binary dependent variable now indicated whether participants chose a cause on the continuous branch or the delayed branch. The posterior mean of the intercept variable was 2.42 (95\% credible interval $[0.33, 4.83]$), indicating that the continuous branch was chosen more often than the delayed branch. The credible intervals for the state variable also excluded 0 (1.39, $[0.32, 2.51]$), suggesting that state sequences led to a stronger preference for the continuous branch than did event sequences.    Credible intervals for both length variables included 0 (medium: 0.17, $[-0.91,1.31]$ ; long: 0.11, $[-1.01,1.21]$, where the reference level is small), suggesting that the preference for the continuous branch was not affected by network length.

@@

%Thus far we have focused on dual branch sequences, and the chain sequences are of interest primarily because they provide additional evidence of a preference for root causes.  NUMBERS HERE

@@

\begin{figure}
\begin{center}
%\includegraphics[width=8cm]{results_E1}
%\includegraphics{figures/expt1.eps}
\includegraphics{figures/expt1test.eps}
\end{center}
\caption{Experiment 1. (a) Order of the four key events for dual branch sequences. The probability raising and continuity accounts both single out \ev{R}{C} as the main cause. (b)  Responses for dual branch sequences. The delayed branch was either inactive (event sequences) or active (state sequences) at the start of the sequence. Error bars show the standard error of the mean estimated using XXX bootstrap samples.} 
\label{fig:expt1}
\end{figure}

\begin{figure}[t]
\begin{center}
%\includegraphics[width=8cm]{results_E2}
\includegraphics{figures/expt2.eps}
%\includegraphics{figures/expt2test.eps}
%\includegraphics{figures/results_E2.eps}
\end{center}
\caption{Experiment 2. (a) The probability raising and continuity accounts make different predictions for dual branch sequences in which \ev{R}{C} occurs before \ev{R}{D}. (b) The two accounts agree when \ev{R}{C} occurs after \ev{R}{D}.
(c) Responses for the three possible orders of the key events.} 
\label{fig:expt2}
\end{figure}

@@

\textbf{Participants}. 100 participants were recruited via Amazon Mechanical Turk and paid \$1 for a 6 minute experiment. Two did not submit a valid completion code, leaving 98 sets of responses for analysis.

@@

\textbf{Results}.  Network orientation (left or right) and position of the curved branch (top or bottom) had no effect and we collapsed across these variables.  Figure~\ref{fig:expt2}c summarizes the results for the three orders listed in Figures \ref{fig:expt2}a and \ref{fig:expt2}b. Consistent with Experiment 1,  \ev{R}{C} is preferred over \ev{R}{D} given the order \{\ev{R}{D},\ev{I}{D},\ev{R}{C}\}, and a similar effect was found for the \{\ev{R}{D},\ev{R}{C},\ev{I}{D}\} sequences. The critical finding is that \ev{R}{C} is also preferred for the \{\ev{R}{C},\ev{R}{D},\ev{I}{D}\} sequences, even though the probability raising account makes the opposite prediction. When referring to the orders we have dropped the fourth event because this event is always \ev{I}{C}.

To further analyze the data we used mixed effects models analogous to the two described for Experiment 1. The first analysis used a logistic regression in which order (i.e.\ the temporal order of the key events) was included as a predictor of a binary dependent variable (root or immediate).  The posterior mean of the intercept variable was 5.75 and the 95\% credible interval $[2.44, 10.52]$, which supports the conclusion that root causes were chosen more often than immediate causes. The credible intervals for both order variables included 0 (\{\ev{R}{D},\ev{R}{C},\ev{I}{D}\}: 2.11, $[-1.53,1.21]$; \{\ev{R}{D},\ev{I}{D},\ev{R}{C}\}: -0.38 $[-1.78,0.90]$, where the reference level was \{\ev{R}{C},\ev{R}{D},\ev{I}{D}\}), suggesting that the preference for root causes was fairly consistent across the three orders.

The second and more critical analysis used a binary dependent variable that indicated whether participants chose a cause on the continuous branch or the delayed branch.  The posterior mean of the intercept variable was 2.0 (95\% credible interval $[1.4, 2.65]$), indicating that the continuous branch was chosen more often than the delayed branch.  The credible intervals for both order variables included 0 ( \{\ev{R}{D},\ev{R}{C},\ev{I}{D},\ev{I}{C}\}: -0.38, $[-0.95,0.21]$ ; \{\ev{R}{D},\ev{I}{D},\ev{R}{C}\}: -0.46, $[-1.03,0.12]$) suggesting that the preference for the continuous branch was of similar strength across all three orders. 
Relative to Experiment 1, the frequency with which \ev{I}{C} is chosen has increased in Experiment 2. This difference may reflect the increased difficulty of Experiment 2. When both branches of a dual branch structure are simultaneously active, keeping track of both root causes and the order in which they occurred is relatively challenging, which may lead some participants to fall back on the simple strategy of choosing the immediate cause that directly precedes the effect.

@@

@@

Because the continuity account considers changes of state in continuous time, it can exploit a ``no coincidences'' principle to identify a single main cause of an effect. This principle applies broadly to causation in physical systems, but is less applicable to social scenarios such as voting scenarios~\cite{livengood13}. Exceptions to the ``no coincidences'' principle are possible even for physical systems, and the continuity account can capture the illusory perceptions of actual causation that arise in some such cases~\cite{thorstadw16}. For example, an 11 year old boy whacked a stick against a telephone pole at the instant at which the Great Northeastern Blackout hit his town, and ran home terrified that he had caused the power outage~\cite{gelbr1965night}.

@@

As suggested earlier, standard formulations of the physical process view treat both \ev{R}{C} and \ev{R}{D} as causes of the effect in Figure~\ref{fig:intro}. Our data are consistent with the possibility that people view \ev{R}{C} and \ev{R}{D} as equally valid causes, but pick \ev{R}{D} when forced to break the tie for reasons that may be relatively superficial. For example, perhaps \ev{R}{D} is preferred because the activation on the continuous branch is continuous both in space and time, and therefore more prototypical of a causal process than the activation along the delay branch. We have planned a future study that asks whether the \ev{R}{C} over \ev{R}{D} is fundamental or superficial using a task in which participants rank multiple candidate causes in order of importance, and in which ties are permitted.

@@

The experimental paradigm we have developed can easily be extended to a more diverse set of causal systems. For example, by introducing preventive causal links, we can explore case of double prevention (e.g.\ cases where an active detector that would have prevented the activation of the GCM is itself inactivated by another detector). \citeA{glymourstone} point out that current work on actual causation is based largely on discussions of an ``infinitesimal fraction'' of the set of possible cases, but our particle detector task can support a comprehensive survey of people's judgments about actual causation.
